---
title: "Regressions logistiques"
author: "Brittany Curry-Sharples"
date: "2024-11-27"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 6
---
## Code
###### Setup
```{r setup, message=FALSE}

library(tidyverse)
```
###### Import Data
```{r import data, message=FALSE}

dataFolder <- './data'                      # sets path to data in r project 

originalFileNames <- list.files(
  path = dataFolder,                        # store all file names in data
  pattern = '\\.csv',                       # folder in a list
  full.names = TRUE)

listNewFileNames <- list()                  # empty list to store new file names

for (file in originalFileNames) {
  
  dataTemp <- read_csv(file)                # import data sets
  
  originalName <- file %>%
    basename() %>%                          # get file name
    str_remove('\\.csv') 
  
  year <- originalName %>% 
    str_extract('\\d{4}([-_]\\d{4})?') %>% # matches year or year range
    str_replace('[-_]', 'to')
  
  if (str_starts(originalName, 'Inuit')) {
    newName <- str_c('dataNorth', year)
  } else if (str_starts(originalName, 'Rest')) {
    newName <- str_c('dataSouth', year)
  } else {
    newName <- originalName                 # if no match, default to original  
  }                                         # name
  
  assign(newName, dataTemp)                # dynamically rename the data set
  
  listNewFileNames <- append(listNewFileNames, newName)
}
```
###### Filter Data
```{r filter data}

variablesToRecode <- list(
  tailleMenage     = c('DHHDGHSZ', 'DHHGHSZ', 'DHHEGHSZ', 'DHHCGHSZ'),
  travaillerEmploi = c('MAC_010', 'GEN_08', 'GENE_08', 'GENC_08', 'GENA_08'),
  santePhysique    = c('GEN_005', 'GEN_01', 'GENE_01', 'GENC_01', 'GENA_01'),
  santeMental      = c('GEN_015', 'GEN_02B', 'GENE_02B', 'GENC_02B'),
  appartCommun     = c('GEN_030', 'GEN_10', 'GENE_10', 'GENC_10', 'GENA_10'),
  typeFumeur       = c('SMK_005', 'SMK_202', 'SMKE_202', 'SMKC_202', 
                       'SMKA_202'),        
  age              = c('DHHGAGE', 'DHHEGAGE', 'DHHCGAGE', 'DHHAGAGE'), 
  typeBuveur       = c('ALCDVTTM', 'ALCDTTM', 'ALCEDTYP', 'ALCCDTYP', 
                       'ALCADTYP'),      
  activitePhys     = c('PAA_030', 'PACDLTI'), 
  activiteSed      = c('SBE_005', 'SACDVTOT', 'SACDTOT', 'SACEDTOT', 'SACCDTOT',  
                       'SACADTOT'),
  satFinanciere    = c('SWL_015'),
  satLogement      = c('SWL_040'),
  relationGens     = c('SPS_050', 'SPS_10'),
  santeNonSatis    = c('UCN_005', 'UCN_010'),
  assuranceMed     = c('INS_005', 'INS_1'),
  revenuMenage     = c('INCDGHH', 'INCGHH', 'INCEGHH', 'INCCGHH', 'INCAGHH'),
  secAlimentaire   = c('FSCDVHF2', 'FSCDVHFS', 'FSCDHFS2', 'FSCDHFS', 
                       'FSCEDHFS', 'FINCF1', 'FINAF1'),
  poids            = c('WTS_M', 'WTSE_M', 'WTSC_M', 'WTSAM')
)

# filter data
for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  
  variablesToKeep <- c()
  
  for (variableList in variablesToRecode) {
    for(variable in variableList) {
      
      if (variable %in% colnames(currentDataset)) {
      
      variablesToKeep <- append(variablesToKeep, variable)
      }
    }
  }
  
  filteredDataset <- currentDataset %>% select(all_of(variablesToKeep))
    
  assign(dataset, filteredDataset)      # keeps standardized name
}

```
###### Recode Variables
```{r recode food (in)security variables}

recode_rules <- list(
  
  FINAF1 = function(x) {
    case_when(
      x == 1 ~ 1,               # secure
      x == 2 ~ 0,               # insecure
      x == 9 ~ NA_integer_,     # ensure NA is of type int  
      is.na(x) ~ NA_integer_,   # handle NA values explicitly
      TRUE ~ as.integer(x)    # force other values to be characters
    )
  },
  
  FINCF1 = function(x) {
    case_when(
      x == 1 ~ 1,
      x == 2 ~ 0,
      x %in% c(6, 9) ~ NA_integer_,
      is.na(x) ~ NA_integer_,  
      TRUE ~ as.integer(x)
    )
  },
  
  FSCDHFS2 = function(x) {
    case_when(
      x == 0 ~ 1,
      x %in% c(1, 2) ~ 0,
      x %in% c(6, 9) ~ NA_integer_,
      TRUE ~ as.integer(x)
    )
  },
  
  FSCDHFS = function(x) {
    case_when(
      x == 0 ~ 1,
      x %in% c(1, 2, 3) ~ 0,
      x == 9 ~ NA_integer_,
      TRUE ~ as.integer(x)
    )
  },
  
  FSCDVHFS = function(x) {
    case_when(
      x == 0 ~ 1,
      x %in% c(1, 2) ~ 0,
      x %in% c(6, 9) ~ NA_integer_,
      TRUE ~ as.integer(x)
    )
  },
  
  FSCDVHF2 = function(x) {
    case_when(
      x == 0 ~ 1,
      x %in% c(1, 2, 3) ~ 0,
      x %in% c(6, 9) ~ NA_integer_,
      TRUE ~ as.integer(x)
    )
  },
  
  FSCEDHFS = function(x) {
    case_when(
      x == 0 ~ 1,
      x %in% c(1, 2, 3) ~ 0,
      x %in% c(6, 9) ~ NA_integer_,
      TRUE ~ as.integer(x)
    )
  }
)

 # Identify the food security variables (secAlimentaire)
  allFoodSecurityVariables <- variablesToRecode$secAlimentaire 

# recode data
for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  #cat("The current dataset is: ", dataset, "\n")
  
  # Identify which food security variable is present in the data set
  foodSecurityVariables <- allFoodSecurityVariables[allFoodSecurityVariables 
                                                    %in%
                                                      colnames(currentDataset)]
    
    #cat("Food security variable found: ")
    #print(foodSecurityVariables)
    #cat("\n")
    
    for (variable in foodSecurityVariables) { # iterate through each FS var (if multiple)
      
      if (variable %in% names(currentDataset)) {
        
      recode_function <- recode_rules[[variable]] # apply recoding rule (see list)
      
      # Recode the food security variable
      currentDataset <- currentDataset %>%
        mutate(!!variable := recode_function(.[[variable]]))
      
      #cat("Recoded", variable, "in ")
      #print(currentDataset) 
      #cat("\n", "\n")
        
      }

  assign(dataset, currentDataset)
    }
}

```
###### Rename Variables
```{r rename variables}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  #cat("\n", "\n")
  #cat("Recoding variables in the dataset: ", dataset, "\n")
  
    for (variable in colnames(currentDataset)) {
      
      originalVariableName <- variable # save original variable name
      
      for (variableType in names(variablesToRecode)) {
        if (originalVariableName %in% variablesToRecode[[variableType]]) {
          
          newVariableName <- str_c(variableType, '(', originalVariableName, ')')
          
          currentDataset <- currentDataset %>% 
            rename(!!newVariableName := !!originalVariableName)
          
          # cat(variable, " ",  variableType, " ", newVariableName, "\n")
          
        }
      }
    }
  assign(dataset, currentDataset)
}
```
###### Logistic Regressions
```{r logistic regressions, warning=TRUE}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  datasetName    <- dataset
  year           <- datasetName %>% 
    str_extract('\\d{4}(to\\d{4})?')  # matches year or year range
  region         <- datasetName %>% 
    str_extract('(North|South)')
  
  #cat("\n", "\n")
  cat("Logistic regression for ", dataset, "\n")
  #cat("----------------------------------------------------", "\n")
  #cat("\n")

  varDep    <- c()
  varsIndep <- c()
  varPoids     <- c()
  
  for (variable in colnames(currentDataset)) {
    
    if (str_detect(variable, '^secAlimentaire')) {
      varDep <- c(varDep, variable)
      
    } else if (str_detect(variable, '^poids')) {
      varPoids <- c(varPoids, currentDataset[[variable]])
      
    } else if (length(unique(currentDataset[[variable]])) > 1) { # ensure variation
      varsIndep <- c(varsIndep, variable)
    }
  }
  #cat("The dependent variable(s) is(are):")
  #cat(varDep, "\n", "\n")
  #cat("The independent variables are:")
  #cat(varsIndep, "\n")
  #cat("the weight variable is: ")
  #cat(varPoids, "\n")

  if (length(varDep) > 1) {
    
    for (varFS in varDep) {
      
      # ensure the dependent variable is an integer
      currentDataset <- currentDataset %>%
        mutate(!!varFS := as.integer(.[[varFS]]))
      
      #cat('the dv is:', varFS)
      #cat(str(varFS), '\n')
      
      cleanedVarDep <- varFS %>% 
        str_remove('secAlimentaire') %>% 
        as.character()
      
      modelName <- str_c('model', region, year, cleanedVarDep)
      
      formula <- as.formula(str_c('`', varFS, '`', '~', 
                                  str_c('`', varsIndep, '`', collapse = ' + ')))
      
      assign(modelName, glm(formula, data = currentDataset, family = binomial, 
                            weights = varPoids))
      
      #cat('Generated models: ', modelName, '\n')
    }
  } else {
    
    modelName <- str_c('model', region, year)
    
    # ensure the dependent variable is an integer
    currentDataset <- currentDataset %>%
    mutate(!!varDep := as.integer(.[[varDep]]))
    
    #cat('the dv is:', varDep)
    #cat(str(varDep), '\n')
    
    formula <- as.formula(str_c('`', varDep, '`', '~', 
                                  str_c('`', varsIndep, '`', collapse = ' + ')))
      
      assign(modelName, glm(formula, data = currentDataset, family = binomial, 
                            weights = varPoids))
      
      #cat('Generated model: ', modelName, "\n")
  }

}

```


## ESCC 2019-2020
### Modèle nord
```{r North 2019-2020, echo=TRUE, warning=TRUE}
formula(modelNorth2019to2020)
summary(modelNorth2019to2020)
```
### Modèle sud
```{r South 2019-2020,echo=TRUE, warning=TRUE}
formula(modelSouth2019to2020)
summary(modelSouth2019to2020)
```

## ESCC 2017-2018
### Modèles nord
```{r North 2017-2018, echo=TRUE, warning=TRUE}
formula(`modelNorth2017to2018(FSCDVHF2)`)
summary(`modelNorth2017to2018(FSCDVHF2)`)

formula(`modelNorth2017to2018(FSCDVHFS)`)
summary(`modelNorth2017to2018(FSCDVHFS)`)
```
### Modèles sud
```{r South 2017-2018, echo=TRUE, warning=TRUE}
formula(`modelSouth2017to2018(FSCDVHF2)`)
summary(`modelSouth2017to2018(FSCDVHF2)`)

formula(`modelSouth2017to2018(FSCDVHFS)`)
summary(`modelSouth2017to2018(FSCDVHFS)`)
```

## ESCC 2015-2016
### Modèle nord
```{r North 2015-2016, echo=TRUE, warning=TRUE}
formula(modelNorth2015to2016)
summary(modelNorth2015to2016)
```
### Modèle sud
```{r South 2015-2016, echo=TRUE, warning=TRUE}
formula(modelSouth2015to2016)
summary(modelSouth2015to2016)
```

## ESCC 2013-2014
### Modèle nord
N. B. (In)sécurité alimentaire non capturé pour les populations du nord.

### Modèle sud
```{r South 2013-2014, echo=TRUE, warning=TRUE}
formula(modelSouth2013to2014)
summary(modelSouth2013to2014)
```

## ESCC 2012
### Modèle nord
```{r North 2012, echo=TRUE, warning=TRUE}
formula(modelNorth2012)
summary(modelNorth2012)
```
### Modèle sud
```{r South 2012, echo=TRUE, warning=TRUE}
formula(modelSouth2012)
summary(modelSouth2012)
```

## ESCC 2009-2010
### Modèle nord
```{r North 2009-2010, echo=TRUE, warning=TRUE}
formula(modelNorth2009to2010)
summary(modelNorth2009to2010)
```
### Modèle sud
```{r South 2009-2010, echo=TRUE, warning=TRUE}
formula(modelSouth2009to2010)
summary(modelSouth2009to2010)
```

## ESCC 2007-2008
### Modèles nord
```{r North 2007-2008, echo=TRUE, warning=TRUE}
formula(`modelNorth2007to2008(FSCDHFS)`)
summary(`modelNorth2007to2008(FSCDHFS)`)

formula(`modelNorth2007to2008(FSCDHFS2)`)
summary(`modelNorth2007to2008(FSCDHFS2)`)
```
### Modèles sud
```{r South 2007-2008, echo=TRUE, warning=TRUE}
formula(`modelSouth2007to2008(FSCDHFS)`)
summary(`modelSouth2007to2008(FSCDHFS)`)

formula(`modelSouth2007to2008(FSCDHFS2)`)
summary(`modelSouth2007to2008(FSCDHFS2)`)
```

## ESCC 2005
### Modèle nord
N. B. (In)sécurité alimentaire non capturé pour les populations du nord.

### Modèle sud
```{r South 2005, echo=TRUE, warning=TRUE}
formula(modelSouth2005)
summary(modelSouth2005)
```

## ESCC 2003
### Modèle nord
```{r North 2003, echo=TRUE, warning=TRUE}
formula(modelNorth2003)
summary(modelNorth2003)
```
### Modèle sud
```{r South 2003, echo=TRUE, warning=TRUE}
formula(modelSouth2003)
summary(modelSouth2003)
```

## ESCC 2001
### Modèle nord
```{r North 2001, echo=TRUE, warning=TRUE}
formula(modelNorth2001)
summary(modelNorth2001)
```
### Modèle sud
```{r South 2001, echo=TRUE, warning=TRUE}
formula(modelSouth2001)
summary(modelSouth2001)
```

