---
title: "Regressions logistiques"
author: "Brittany Curry-Sharples"
date: "2024-11-27"
output:
  pdf_document:
    toc: true
    toc_depth: '6'
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 6
---
## Code
###### Setup
```{r setup, message=FALSE}

library(tidyverse)
```
###### Import Data
```{r import data, message=FALSE}

dataFolder <- './data'                      # sets path to data in r project 

originalFileNames <- list.files(
  path = dataFolder,                        # store all file names in data
  pattern = '\\.csv',                       # folder in a list
  full.names = TRUE)

listNewFileNames <- list()                  # empty list to store new file names

for (file in originalFileNames) {
  
  dataTemp <- read_csv(file)                # import data sets
  
  originalName <- file %>%
    basename() %>%                          # get file name
    str_remove('\\.csv') 
  
  year <- originalName %>% 
    str_extract('\\d{4}([-_]\\d{4})?') %>% # matches year or year range
    str_replace('[-_]', 'to')
  
  if (str_starts(originalName, 'Inuit')) {
    newName <- str_c('dataNorth', year)
  } else if (str_starts(originalName, 'Rest')) {
    newName <- str_c('dataSouth', year)
  } else {
    newName <- originalName                 # if no match, default to original  
  }                                         # name
  
  assign(newName, dataTemp)                # dynamically rename the data set
  
  listNewFileNames <- append(listNewFileNames, newName)
}
```
###### Filter Data
```{r filter data}

variablesToRecode <- list(
  id               = c('ADM_RNO1', 'ADM_RNO', 'ADME_RNO', 'ADMC_RNO', 
                       'ADMA_RNO'),
  tailleMenage     = c('DHHDGHSZ', 'DHHGHSZ', 'DHHEGHSZ', 'DHHCGHSZ'),
  travaillerEmploi = c('MAC_010', 'GEN_08', 'GENE_08', 'GENC_08', 'GENA_08'),
  santePhysique    = c('GEN_005', 'GEN_01', 'GENE_01', 'GENC_01', 'GENA_01'),
  santeMental      = c('GEN_015', 'GEN_02B', 'GENE_02B', 'GENC_02B'),
  appartCommun     = c('GEN_030', 'GEN_10', 'GENE_10', 'GENC_10', 'GENA_10'),
  typeFumeur       = c('SMK_005', 'SMK_202', 'SMKE_202', 'SMKC_202', 
                       'SMKA_202'),        
  age              = c('DHHGAGE', 'DHHEGAGE', 'DHHCGAGE', 'DHHAGAGE'), 
  typeBuveur       = c('ALCDVTTM', 'ALCDTTM', 'ALCEDTYP', 'ALCCDTYP', 
                       'ALCADTYP'),      
  #activitePhys     = c('PAA_030', 'PACDLTI'), 
  #activiteSed      = c('SBE_005', 'SACDVTOT', 'SACDTOT', 'SACEDTOT', 'SACCDTOT',  
  #                     'SACADTOT'),
  #satFinanciere    = c('SWL_015'),
  #satLogement      = c('SWL_040'),
  #relationGens     = c('SPS_050', 'SPS_10'),
  #santeNonSatis    = c('UCN_005', 'UCN_010'),
  #assuranceMed     = c('INS_005', 'INS_1'),
  revenuMenage     = c('INCDGHH', 'INCGHH', 'INCEGHH', 'INCCGHH', 'INCAGHH'),
  secAlimentaire   = c('FSCDVHF2', 'FSCDVHFS', 'FSCDHFS2', 'FSCDHFS', 
                       'FSCEDHFS', 'FINCF1', 'FINAF1'),
  poids            = c('WTS_M', 'WTSE_M', 'WTSC_M', 'WTSAM')
)

# filter data
for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  
  variablesToKeep <- c()
  
  for (variableList in variablesToRecode) {
    for(variable in variableList) {
      
      if (variable %in% colnames(currentDataset)) {
      
      variablesToKeep <- append(variablesToKeep, variable)
      }
    }
  }
  
  filteredDataset <- currentDataset %>% select(all_of(variablesToKeep))
    
  assign(dataset, filteredDataset)      # keeps standardized name
}

# ensure ID values are unique
for (dataset in listNewFileNames) {

   currentDataset <- get(dataset)
   cat('\n')
   cat('Checking dataset: ', dataset, '\n')

   for (element in variablesToRecode[['id']]) {

     if (element %in% names(currentDataset)) {

       check <- any(duplicated(currentDataset[[element]]))  # return TRUE if duplicates, false otherwise

       cat('Is any element in', element, 'duplicated?: ', check, '\n')

     }
   }
}

```
###### Append Survey Year to Variables
``` {r}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  
    year <- dataset %>% 
      str_extract('\\d{4}(to\\d{4})?')  # matches year or year range
    
    for (variable in names(currentDataset)) {
      
      originalVariableName <- variable
      
      newVariableName <- str_c(originalVariableName, '_', year)
      
      currentDataset <- currentDataset %>%
        rename(!!newVariableName := !!originalVariableName)
    }
    assign(dataset, currentDataset)
}

```
###### Recode Variables
```{r recode food (in)security variables}

recode_rules <- list( # cycles 2013-2014 and 2005 not recoded because there is no data for FS for Northern regions, 2003 and 2001 excluded because variables differ from other years (esp. fs, age, drinking status)

    # food (in)security
  
  FSCDVHF2_2019to2020 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2, 3) ~ 'Insécurité alimentaire',
      x %in% c(6, 9,                          # ensure NA is of type character
               is.na(x)) ~ NA_character_,     # and handle NA values explicitly 
      TRUE ~ as.character(x)                  # force other values to be char
    )
  },
  
   FSCDVHF2_2017to2018 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2, 3) ~ 'Insécurité alimentaire',
      x %in% c(9, is.na(x)) ~ NA_character_,     
      TRUE ~ as.character(x)                 
    )
  },
  
  FSCDVHFS_2017to2018 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2) ~ 'Insécurité alimentaire',
      x %in% c(9, is.na(x)) ~ NA_character_,     
      TRUE ~ as.character(x)                 
    )
  },
  
  FSCDVHFS_2015to2016 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2) ~ 'Insécurité alimentaire',
      x %in% c(6, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
    FSCDHFS2_2012 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2) ~ 'Insécurité alimentaire',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  FSCDHFS2_2009to2010 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2) ~ 'Insécurité alimentaire',
      x %in% c(6, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  FSCDHFS2_2007to2008 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2) ~ 'Insécurité alimentaire',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  FSCDHFS_2007to2008 = function(x) {
    case_when(
      x == 0 ~ 'Sécurité alimentaire',
      x %in% c(1, 2, 3) ~ 'Insécurité alimentaire',
      x %in% c(9, is.na(x)) ~ NA_character_,     
      TRUE ~ as.character(x)                 
    )
  },
  
  # FINCF1_2003 = function(x) {
  #   case_when(
  #     x == 2 ~ 'Sécurité alimentaire',
  #     x == 1 ~ 'Insécurité alimentaire',
  #     x %in% c(6, 9, is.na(x)) ~ NA_character_,     
  #     TRUE ~ as.character(x)                 
  #   )
  # },
  
  # household size
  
  DHHDGHSZ_2019to2020 = function(x) {
    case_when(
      x == 1 ~ 'Vit seul (taille du ménage=1)',
      x == 2 ~ 'Taille du ménage 2 ou plus',
      x %in% c(9, is.na(x)) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },  
    
  DHHDGHSZ_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Vit seul (taille du ménage=1)',
      x %in% c(2, 3, 4, 5) ~ 'Taille du ménage 2 ou plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  }, 
  
  DHHDGHSZ_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Vit seul (taille du ménage=1)',
      x %in% c(2, 3, 4, 5) ~ 'Taille du ménage 2 ou plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )   
  },
  
  DHHGHSZ_2012 = function(x) {
    case_when(
      x == 1 ~ 'Vit seul (taille du ménage=1)',
      x %in% c(2, 3, 4, 5) ~ 'Taille du ménage 2 ou plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  DHHGHSZ_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Vit seul (taille du ménage=1)',
      x %in% c(2, 3, 4, 5) ~ 'Taille du ménage 2 ou plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  DHHGHSZ_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Vit seul (taille du ménage=1)',
      x %in% c(2, 3, 4, 5) ~ 'Taille du ménage 2 ou plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  # DHHCGHSZ_2003 = function(x) {
  #   case_when(
  #     x == 1 ~ 'Vit seul (taille du ménage=1)',
  #     x %in% c(2, 3, 4, 5) ~ 'Taille du ménage 2 ou plus',
  #     x == is.na(x) ~ NA_character_, 
  #     TRUE ~ as.character(x)
  #   )
  # },
  
  # employment status
  
  MAC_010_2019to2020 = function(x) {
    case_when(
      x == 1 ~ 'Oui',
      x == 2 ~ 'Non', 
      x %in% c(6, 7, 8, 9, is.na(x)) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },
  
  MAC_010_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Oui',
      x == 2 ~ 'Non', 
      x %in% c(6, 7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  MAC_010_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Oui',
      x == 2 ~ 'Non', 
      x %in% c(6, 7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  GEN_08_2012 = function(x) {
    case_when(
      x == 1 ~ 'Oui',
      x == 2 ~ 'Non', 
      x %in% c(6, 7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  GEN_08_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Oui',
      x == 2 ~ 'Non', 
      x %in% c(6, 7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  GEN_08_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Oui',
      x == 2 ~ 'Non', 
      x %in% c(6, 7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )
  },
  
  # GENC_08_2003 = function(x) {
  #   case_when(
  #     x == 1 ~ 'Oui',
  #     x == 2 ~ 'Non', 
  #     x %in% c(6, 7, 8, 9, is.na(x)) ~ NA_character_, 
  #     TRUE ~ as.character(x)
  #   )
  # },
  
  # Perceived physical health 

  GEN_005_2019to2020 = function(x) {
    case_when( 
      x == 1 ~ 'Excellente', 
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable', 
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, is.na(x)) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },
  
  GEN_005_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
    
  GEN_005_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  GEN_01_2012 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  GEN_01_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  GEN_01_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )   
  },
  
  # GENC_01_2003 = function(x) {
  #   case_when(
  #     x == 1 ~ 'Excellente',
  #     x == 2 ~ 'Très bonne', 
  #     x == 3 ~ 'Bonne', 
  #     x == 4 ~ 'Passable',
  #     x == 5 ~ 'Mauvaise',
  #     x %in% c(7, 8, is.na(x)) ~ NA_character_, 
  #     TRUE ~ as.character(x)
  #   ) 
  # },
  
  # Perceived mental health

  GEN_015_2019to2020 = function(x) {
    case_when(
      x == 1 ~ 'Excellente', 
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_,
     TRUE ~ as.character(x)
    )
  },
  
  GEN_015_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  GEN_015_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  GEN_02B_2012 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  GEN_02B_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  GEN_02B_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Excellente',
      x == 2 ~ 'Très bonne', 
      x == 3 ~ 'Bonne', 
      x == 4 ~ 'Passable',
      x == 5 ~ 'Mauvaise',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  # GENC_02B_2003 = function(x) {
  #   case_when(
  #     x == 1 ~ 'Excellente',
  #     x == 2 ~ 'Très bonne', 
  #     x == 3 ~ 'Bonne', 
  #     x == 4 ~ 'Passable',
  #     x == 5 ~ 'Mauvaise',
  #     x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
  #     TRUE ~ as.character(x)
  #   ) 
  # },
  
  # Community belonging

  GEN_030_2019to2020 = function(x) {
    case_when(
      x == 1 ~ 'Très fort',
      x == 2 ~ 'Plutôt fort', 
      x == 3 ~ 'Plutôt faible', 
      x == 4 ~ 'Très faible', 
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },
  
  GEN_030_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Très fort',
      x == 2 ~ 'Plutôt fort', 
      x == 3 ~ 'Plutôt faible', 
      x == 4 ~ 'Très faible',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  GEN_030_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Très fort',
      x == 2 ~ 'Plutôt fort', 
      x == 3 ~ 'Plutôt faible', 
      x == 4 ~ 'Très faible',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  GEN_10_2012 = function(x) {
    case_when(
      x == 1 ~ 'Très fort',
      x == 2 ~ 'Plutôt fort', 
      x == 3 ~ 'Plutôt faible', 
      x == 4 ~ 'Très faible',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  GEN_10_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Très fort',
      x == 2 ~ 'Plutôt fort', 
      x == 3 ~ 'Plutôt faible', 
      x == 4 ~ 'Très faible',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  GEN_10_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Très fort',
      x == 2 ~ 'Plutôt fort', 
      x == 3 ~ 'Plutôt faible', 
      x == 4 ~ 'Très faible',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  # GENC_10_2003 = function(x) {
  #   case_when(
  #     x == 1 ~ 'Très fort',
  #     x == 2 ~ 'Plutôt fort', 
  #     x == 3 ~ 'Plutôt faible', 
  #     x == 4 ~ 'Très faible',
  #     x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
  #     TRUE ~ as.character(x)
  #   ) 
  # },
  
  # Smoking status

  SMK_005_2019to2020 = function(x) {
    case_when(
      x == 1 ~ 'Tous les jours',
      x == 2 ~ 'À l’occasion',
      x == 3 ~ 'Jamais',
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },
  
  SMK_005_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Tous les jours',
      x == 2 ~ 'À l’occasion', 
      x == 3 ~ 'Jamais', 
      x %in% c(7, 8, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  }, 
  
  SMK_005_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Tous les jours',
      x == 2 ~ 'À l’occasion', 
      x == 3 ~ 'Jamais', 
      x %in% c(7, 8, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  SMK_202_2012 = function(x) {
    case_when(
      x == 1 ~ 'Tous les jours',
      x == 2 ~ 'À l’occasion', 
      x == 3 ~ 'Jamais', 
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  SMK_202_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Tous les jours',
      x == 2 ~ 'À l’occasion', 
      x == 3 ~ 'Jamais', 
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  SMK_202_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Tous les jours',
      x == 2 ~ 'À l’occasion', 
      x == 3 ~ 'Jamais', 
      x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  # SMKC_202_2003 = function(x) {
  #   case_when(
  #     x == 1 ~ 'Tous les jours',
  #     x == 2 ~ 'À l’occasion', 
  #     x == 3 ~ 'Jamais', 
  #     x %in% c(7, 8, 9, is.na(x)) ~ NA_character_, 
  #     TRUE ~ as.character(x)
  #   ) 
  # },
  
  # Age group

  DHHGAGE_2019to2020 = function(x) {
    case_when(
      x == 1 ~ '12 à 17 ans', 
      x == 2 ~ '18 à 34 ans', 
      x == 3 ~ '35 à 49 ans',
      x == 4 ~ '50 à 64 ans',
      x == 5 ~ '65 ans et plus',
      x == is.na(x) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },
  
  DHHGAGE_2017to2018 = function(x) {
    case_when(
      x %in% c(1, 2) ~ '12 à 17 ans',
      x %in% c(3, 4, 5, 6) ~ '18 à 34 ans', 
      x %in% c(7, 8, 9) ~ '35 à 49 ans',
      x %in% c(10, 11, 12) ~ '50 à 64 ans',
      x %in% c(13, 14, 15, 16) ~ '65 ans et plus',
      x == is.na(x) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  DHHGAGE_2015to2016 = function(x) {
    case_when(
      x %in% c(1, 2) ~ '12 à 17 ans',
      x %in% c(3, 4, 5, 6) ~ '18 à 34 ans', 
      x %in% c(7, 8, 9) ~ '35 à 49 ans',
      x %in% c(10, 11, 12) ~ '50 à 64 ans',
      x %in% c(13, 14, 15, 16) ~ '65 ans et plus',
      x == is.na(x) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  DHHGAGE_2012 = function(x) {
    case_when(
      x %in% c(1, 2) ~ '12 à 17 ans',
      x %in% c(3, 4, 5, 6) ~ '18 à 34 ans', 
      x %in% c(7, 8, 9) ~ '35 à 49 ans',
      x %in% c(10, 11, 12) ~ '50 à 64 ans',
      x %in% c(13, 14, 15, 16) ~ '65 ans et plus',
      x == is.na(x) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  DHHGAGE_2009to2010 = function(x) {
    case_when(
      x %in% c(1, 2) ~ '12 à 17 ans',
      x %in% c(3, 4, 5, 6) ~ '18 à 34 ans', 
      x %in% c(7, 8, 9) ~ '35 à 49 ans',
      x %in% c(10, 11, 12) ~ '50 à 64 ans',
      x %in% c(13, 14, 15, 16) ~ '65 ans et plus',
      x == is.na(x) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  DHHGAGE_2007to2008 = function(x) {
    case_when(
      x %in% c(1, 2) ~ '12 à 17 ans',
      x %in% c(3, 4, 5, 6) ~ '18 à 34 ans', 
      x %in% c(7, 8, 9) ~ '35 à 49 ans',
      x %in% c(10, 11, 12) ~ '50 à 64 ans',
      x %in% c(13, 14, 15, 16) ~ '65 ans et plus',
      x == is.na(x) ~ NA_character_, 
      TRUE ~ as.character(x)
    )  
  },
  
  # Alcohol consumption status

  ALCDVTTM_2019to2020 = function(x) {
    case_when(
      x == 1 ~ 'Buveur régulier',
      x == 2 ~ 'Buveur occasionnel',
      x == 3 ~ 'N’a pas bu au cours des 12 derniers mois', 
      x %in% c(9, is.na(x)) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },
  
    ALCDVTTM_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Buveur régulier',
      x == 2 ~ 'Buveur occasionnel', 
      x == 3 ~ 'N’a pas bu au cours des 12 derniers mois', 
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
    ALCDVTTM_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Buveur régulier',
      x == 2 ~ 'Buveur occasionnel', 
      x == 3 ~ 'N’a pas bu au cours des 12 derniers mois', 
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
  ALCDTTM_2012 = function(x) {
    case_when(
      x == 1 ~ 'Buveur régulier',
      x == 2 ~ 'Buveur occasionnel', 
      x == 3 ~ 'N’a pas bu au cours des 12 derniers mois', 
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  ALCDTTM_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Buveur régulier',
      x == 2 ~ 'Buveur occasionnel', 
      x == 3 ~ 'N’a pas bu au cours des 12 derniers mois', 
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  ALCDTTM_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Buveur régulier',
      x == 2 ~ 'Buveur occasionnel', 
      x == 3 ~ 'N’a pas bu au cours des 12 derniers mois', 
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  # Household income

  INCDGHH_2019to2020 = function(x) {
    case_when(
      x == 1 ~ 'Aucun revenu ou moins de 20 000 $',
      x == 2 ~ 'De 20 000 $ à 39 999 $',
      x == 3 ~ 'De 40 000 $ à 59 999 $',
      x == 4 ~ 'De 60 000 $ à 79 999 $',
      x == 5 ~ '80 000 $ et plus',
      x %in% c(9, is.na(x)) ~ NA_character_,
      TRUE ~ as.character(x)
    )
  },
  
  INCDGHH_2017to2018 = function(x) {
    case_when(
      x == 1 ~ 'Aucun revenu ou moins de 20 000 $',
      x == 2 ~ 'De 20 000 $ à 39 999 $', 
      x == 3 ~ 'De 40 000 $ à 59 999 $', 
      x == 4 ~ 'De 60 000 $ à 79 999 $',
      x == 5 ~ '80 000 $ et plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  },
  
    INCDGHH_2015to2016 = function(x) {
    case_when(
      x == 1 ~ 'Aucun revenu ou moins de 20 000 $',
      x == 2 ~ 'De 20 000 $ à 39 999 $', 
      x == 3 ~ 'De 40 000 $ à 59 999 $', 
      x == 4 ~ 'De 60 000 $ à 79 999 $',
      x == 5 ~ '80 000 $ et plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )    
  }, 
  
  INCGHH_2012 = function(x) {
    case_when(
      x == 1 ~ 'Aucun revenu ou moins de 20 000 $',
      x == 2 ~ 'De 20 000 $ à 39 999 $', 
      x == 3 ~ 'De 40 000 $ à 59 999 $', 
      x == 4 ~ 'De 60 000 $ à 79 999 $',
      x == 5 ~ '80 000 $ et plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    )   
  },
  
  INCGHH_2009to2010 = function(x) {
    case_when(
      x == 1 ~ 'Aucun revenu ou moins de 20 000 $',
      x == 2 ~ 'De 20 000 $ à 39 999 $', 
      x == 3 ~ 'De 40 000 $ à 59 999 $', 
      x == 4 ~ 'De 60 000 $ à 79 999 $',
      x == 5 ~ '80 000 $ et plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  },
  
  INCGHH_2007to2008 = function(x) {
    case_when(
      x == 1 ~ 'Aucun revenu ou moins de 20 000 $',
      x == 2 ~ 'De 20 000 $ à 39 999 $', 
      x == 3 ~ 'De 40 000 $ à 59 999 $', 
      x == 4 ~ 'De 60 000 $ à 79 999 $',
      x == 5 ~ '80 000 $ et plus',
      x %in% c(9, is.na(x)) ~ NA_character_, 
      TRUE ~ as.character(x)
    ) 
  }
  
)

 # Identify the food security variables (secAlimentaire)
  allFoodSecurityVariables <- variablesToRecode$secAlimentaire 

# recode data
for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  #cat("The current dataset is: ", dataset, "\n")
    
    for (variable in names(recode_rules)) { # iterate through each var to be recoded
      
      if (variable %in% names(currentDataset)) {
        
      recode_function <- recode_rules[[variable]] # apply recoding rule (see list)
      
      # Recode the food security variable
      currentDataset <- currentDataset %>%
        mutate(!!variable := recode_function(.[[variable]]))
      
      #cat("Recoded", variable, "in ")
      #print(currentDataset) 
      #cat("\n", "\n")
        
      }

  assign(dataset, currentDataset)
    }
}

```
###### Rename Variables
```{r rename variables}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  #cat("\n", "\n")
  #cat("Recoding variables in the dataset: ", dataset, "\n")
  
    for (variable in colnames(currentDataset)) {
      
      originalVariableName <- variable %>%  # save original variable name
        str_remove('_\\d{4}(to\\d{4})?')
      
      #cat('[debug]: Original column:', variable, '-> After str_remove:', 
          #originalVariableName, '\n')
      
      #cat('[debug]: Columns in dataset:', colnames(currentDataset), '\n')
      
      for (variableType in names(variablesToRecode)) {
        
        #cat('[debug]: Current variableType: ', variableType, '\n')
        
         #cat('[debug]: checking if ', variable, 'is in variablesToRecode', '\n')
        
        if (originalVariableName %in% variablesToRecode[[variableType]]) {
          
          newVariableName <- str_c(variableType, '(', originalVariableName, ')')
          
          #cat('[debug]: New variable name: ', newVariableName, '\n', '\n')
          
          currentDataset <- currentDataset %>% 
            rename(!!newVariableName := !!variable) # use `variable` here bc in dataset the years are still appended
          
          # cat(variable, " ",  variableType, " ", newVariableName, "\n")
          
        }
      }
    }
  assign(dataset, currentDataset)
}
```
###### Reshape Data
```{r reshape data}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  
  varId <- names(currentDataset) %>% 
    str_extract('^id\\([A-Za-z0-9_]+\\)') %>% 
    {names(currentDataset)[which(!is.na(.))]} %>%
    unique()
    
  #print(varId)

  varPoids <- names(currentDataset) %>% 
    str_extract('^poids\\([A-Za-z0-9_]+\\)') %>% 
    {names(currentDataset)[which(!is.na(.))]} %>%
    unique()
  
  #print(varPoids)
  
  varDep <- names(currentDataset) %>% 
    str_extract('^secAlimentaire\\([A-Za-z0-9_]+\\)') %>% 
    {names(currentDataset)[which(!is.na(.))]} %>%
    unique()
  
  #print(varDep)

  dataReshaped <- currentDataset %>%
    mutate(across(everything(), ~ if_else(is.na(.), 'NA', as.character(.)))) %>%   # replace NA values with the character 'NA' and cast all other values into characters bc didn't recode for years 2001-2012
    pivot_longer(cols = -all_of(c(varId, varPoids, varDep)),
                 names_to = 'variable',
                 values_to = 'values'
                 ) %>%
    mutate(
      across(starts_with('secAlimentaire'), ~ case_when(
        . == 'Insécurité alimentaire' ~ '0',
        . == 'Sécurité alimentaire' ~ '1',
        . == 'NA' ~ 'NA',
        is.na(.) ~ NA_character_, # keep NA
        TRUE ~ . # in case there's an unexpected value, keep the original (important for years not recoded)
        )),
      newColumn = str_c(variable, values, sep = '_'),
      values = if_else(values == 'NA', 'NA', '1')# Only replace non-NA values with "1"
      ) %>%
    pivot_wider(id_cols = c(all_of(varId), all_of(varPoids), all_of(varDep)),
                names_from = newColumn,
                values_from = values, values_fill = '0') %>%
    select(order(names(.))) %>%   # Alphabetical order
    select(all_of(varId), all_of(varPoids), all_of(varDep), everything()) %>%    # Order columns alphabetically and move 'id' to the front
    select(-ends_with('_NA')) # remove all NA columns
  
  assign(dataset, dataReshaped)
}

```
###### Delete Rows with NA Values for Food (In)Security

```{r filter NA FS values out}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  
  dataFiltered <- currentDataset %>% 
    filter(if_any(starts_with('secAlimentaire'), ~ . != 'NA'))
  
  assign(dataset, dataFiltered)
}

```

###### Ensure Datasets are Numeric Type
```{r numeric type}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  
  dataNumeric <- currentDataset %>% 
    mutate(across(everything(), ~ case_when(
       . == 'NA' ~ NA_real_, # replace 'NA' string with numeric NA 
        is.na(.) ~ NA_real_, # if any NA values cast as numeric
        TRUE ~ as.numeric(.) # else cast to numeric
    )))
  
    assign(dataset, dataNumeric)
}
```
###### Export Cleaned Datasets
```{r export clean data}
for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  
  folderPath <- file.path(getwd(), 'cleanData') 
  
  filename <- file.path(folderPath, str_c(dataset, '.csv'))
  
  write.csv(currentDataset, file = filename, row.names = FALSE)
}
```

###### Logistic Regressions
```{r logistic regressions, warning=TRUE}

for (dataset in listNewFileNames) {
  
  currentDataset <- get(dataset)
  datasetName    <- dataset
  year           <- datasetName %>% 
    str_extract('\\d{4}(to\\d{4})?')  # matches year or year range
  region         <- datasetName %>% 
    str_extract('(North|South)')
  
  cat("\n", "\n")
  cat("Logistic regression for ", dataset, "\n")
  cat("----------------------------------------------------", "\n")
  cat("\n")
  
  if (year == '2015to2016' | year == '2017to2018' | year == '2019to2020') {
    
    varId <- c()
     varDep    <- c()
  varsIndep <- c()
  varPoids     <- c()
  
  for (variable in colnames(currentDataset)) {
    
    if (str_detect(variable, '^secAlimentaire')) {
      varDep <- c(varDep, variable)
      
    } else if (str_detect(variable, '^poids')) {
      varPoids <- c(varPoids, currentDataset[[variable]])
      
    } else if (str_detect(variable, '^id')) {
      varId <- c(varId, currentDataset[[variable]]) # make sure ID is not included in model
      
    } else if (length(unique(currentDataset[[variable]])) > 1) { # ensure variation
      varsIndep <- c(varsIndep, variable)
    } 
  }
  #cat("The dependent variable(s) is(are):")
  #cat(varDep, "\n", "\n")
  #cat("The independent variables are:")
  #cat(varsIndep, "\n")
  #cat("the weight variable is: ")
  #cat(varPoids, "\n")

  if (length(varDep) > 1) {
    
    for (varFS in varDep) {
      
      # ensure the dependent variable is an integer
      currentDataset <- currentDataset %>%
        mutate(!!varFS := as.integer(.[[varFS]]))
      
      #cat('the dv is:', varFS)
      #cat(str(varFS), '\n')
      
      cleanedVarDep <- varFS %>% 
        str_remove('secAlimentaire') %>% 
        as.character()
      
      modelName <- str_c('model', region, year, cleanedVarDep)
      
      formula <- as.formula(str_c('`', varFS, '`', '~', 
                                  str_c('`', varsIndep, '`', collapse = ' + ')))
      
      assign(modelName, glm(formula, data = currentDataset, family = binomial, 
                            weights = varPoids))
      
      #cat('Generated models: ', modelName, '\n')
    }
  } else {
    
    modelName <- str_c('model', region, year)
    
    # ensure the dependent variable is an integer
    currentDataset <- currentDataset %>%
    mutate(!!varDep := as.integer(.[[varDep]]))
    
    #cat('the dv is:', varDep)
    #cat(str(varDep), '\n')
    
    formula <- as.formula(str_c('`', varDep, '`', '~', 
                                  str_c('`', varsIndep, '`', collapse = ' + ')))
      
      assign(modelName, glm(formula, data = currentDataset, family = binomial, 
                            weights = varPoids))
      
      #cat('Generated model: ', modelName, "\n")
    }
  }
}

```


## ESCC 2019-2020
### Modèle nord
```{r North 2019-2020, echo=TRUE, warning=TRUE}
head(dataNorth2019to2020)
formula(modelNorth2019to2020)
summary(modelNorth2019to2020)
```
### Modèle sud
```{r South 2019-2020,echo=TRUE, warning=TRUE}
head(dataSouth2019to2020)
formula(modelSouth2019to2020)
summary(modelSouth2019to2020)
```

## ESCC 2017-2018
### Modèles nord
```{r North 2017-2018, echo=TRUE, warning=TRUE}
formula(`modelNorth2017to2018(FSCDVHF2)`)
summary(`modelNorth2017to2018(FSCDVHF2)`)

formula(`modelNorth2017to2018(FSCDVHFS)`)
summary(`modelNorth2017to2018(FSCDVHFS)`)
```
### Modèles sud
```{r South 2017-2018, echo=TRUE, warning=TRUE}
formula(`modelSouth2017to2018(FSCDVHF2)`)
summary(`modelSouth2017to2018(FSCDVHF2)`)

formula(`modelSouth2017to2018(FSCDVHFS)`)
summary(`modelSouth2017to2018(FSCDVHFS)`)
```

## ESCC 2015-2016
### Modèle nord
```{r North 2015-2016, echo=TRUE, warning=TRUE}
formula(modelNorth2015to2016)
summary(modelNorth2015to2016)
```
### Modèle sud
```{r South 2015-2016, echo=TRUE, warning=TRUE}
formula(modelSouth2015to2016)
summary(modelSouth2015to2016)
```

## ESCC 2013-2014
### Modèle nord
N. B. (In)sécurité alimentaire non capturé pour les populations du nord.

### Modèle sud
```{r South 2013-2014, echo=TRUE, warning=TRUE}
#formula(modelSouth2013to2014)
#summary(modelSouth2013to2014)
```

## ESCC 2012
### Modèle nord
```{r North 2012, echo=TRUE, warning=TRUE}
#formula(modelNorth2012)
#summary(modelNorth2012)
```
### Modèle sud
```{r South 2012, echo=TRUE, warning=TRUE}
#formula(modelSouth2012)
#summary(modelSouth2012)
```

## ESCC 2009-2010
### Modèle nord
```{r North 2009-2010, echo=TRUE, warning=TRUE}
#formula(modelNorth2009to2010)
#summary(modelNorth2009to2010)
```
### Modèle sud
```{r South 2009-2010, echo=TRUE, warning=TRUE}
#formula(modelSouth2009to2010)
#summary(modelSouth2009to2010)
```

## ESCC 2007-2008
### Modèles nord
```{r North 2007-2008, echo=TRUE, warning=TRUE}
#formula(`modelNorth2007to2008(FSCDHFS)`)
#summary(`modelNorth2007to2008(FSCDHFS)`)

#formula(`modelNorth2007to2008(FSCDHFS2)`)
#summary(`modelNorth2007to2008(FSCDHFS2)`)
```
### Modèles sud
```{r South 2007-2008, echo=TRUE, warning=TRUE}
#formula(`modelSouth2007to2008(FSCDHFS)`)
#summary(`modelSouth2007to2008(FSCDHFS)`)

#formula(`modelSouth2007to2008(FSCDHFS2)`)
#summary(`modelSouth2007to2008(FSCDHFS2)`)
```

## ESCC 2005
### Modèle nord
N. B. (In)sécurité alimentaire non capturé pour les populations du nord.

### Modèle sud
```{r South 2005, echo=TRUE, warning=TRUE}
#formula(modelSouth2005)
#summary(modelSouth2005)
```

## ESCC 2003
### Modèle nord
```{r North 2003, echo=TRUE, warning=TRUE}
#formula(modelNorth2003)
#summary(modelNorth2003)
```
### Modèle sud
```{r South 2003, echo=TRUE, warning=TRUE}
#formula(modelSouth2003)
#summary(modelSouth2003)
```

## ESCC 2001
### Modèle nord
```{r North 2001, echo=TRUE, warning=TRUE}
#formula(modelNorth2001)
#summary(modelNorth2001)
```
### Modèle sud
```{r South 2001, echo=TRUE, warning=TRUE}
#formula(modelSouth2001)
#summary(modelSouth2001)
```

